Assessing the Global COVID19 Impact on Air Transport with Open Data

### 2021-07-12 paper.Rmd
network level assessment, combined timeline (need to work out nicer design)

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(stats)
library(zoo)

@@

```{r}
# load analytic dataset
us <- ds %>% 
  filter(ADEP_REG == "US" | ADES_REG == "US") %>%
  group_by(YR = year(DATE), DATE) %>% 
  summarize(N = n(), .groups = "drop")

# static thresholds, x-th percentile rounded to lower integer
pcts   <- c(0.8, 0.6, 0.3, 0.1)
us_ref <- us %>% filter(YR == 2019) %>%
  pull(N) %>% quantile(probs = pcts) %>% floor()

us_ref2 <- us %>% filter(YR == 2019, DATE >= lubridate::ymd("2019-07-01")) %>%
  pull(N) %>% quantile(probs = pcts) %>% floor()

max_N <- max(us$N)
lvls  <- c("First", "Second", "Third", "Fourth", "Fifth")
us <- us %>% 
  mutate( THRES  = cut(N, breaks = c(max_N, us_ref, 0),  labels = lvls)
         ,THRES2 = cut(N, breaks = c(max_N, us_ref2, 0), labels = lvls)
         )
```

@@

```{r fig-us-timeline, fig.cap="(ref:fig-us-timeline)"}
us %>%
  ggplot(aes(x = DATE)) +
  geom_hline(yintercept = us_ref2, linetype = "dotdash") +
  geom_point(aes(y = N, color = THRES2), alpha = 0.3) +
  geom_line( aes(y = N, colour = THRES2, group = 1)) +
  theme_minimal()
```

@@

1.1

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(stats)
library(zoo)
library(patchwork)

@@

```{r}
# summarise regional level
## -------- percentile thresholds
pcts   <- c(0.8, 0.6, 0.3, 0.1)

## -------- function definition
regional_timeline <- function(.ds, .reg, .pcts=pcts){
us <- .ds %>% 
  filter(ADEP_REG == .reg | ADES_REG == .reg) %>%
  group_by(YR = year(DATE), DATE) %>% 
  summarize(N = n(), .groups = "drop")

# static thresholds, x-th percentile rounded to lower integer
us_ref <- us %>% filter(YR == 2019) %>%
  pull(N) %>% quantile(probs = .pcts) %>% floor()

us_ref2 <- us %>% filter(YR == 2019, DATE >= lubridate::ymd("2019-07-01")) %>%
  pull(N) %>% quantile(probs = .pcts) %>% floor()

max_N <- max(us$N)
lvls  <- c("First", "Second", "Third", "Fourth", "Fifth")
us <- us %>% 
  mutate( THRES  = cut(N, breaks = c(max_N, us_ref, 0),  labels = lvls)
         ,THRES2 = cut(N, breaks = c(max_N, us_ref2, 0), labels = lvls)
         )
return(list(us, us_ref, us_ref2))
}

## do for us
res <- regional_timeline(ds, "US")
us  <- res[[1]]
us_ref2 <- res[[3]]
```

@@

```{r fig-us-timeline, fig.cap="(ref:fig-us-timeline)"}
plot_timeline <- function(.df, .pct_ref){
 p <- .df %>%
  ggplot(aes(x = DATE)) +
  geom_hline(yintercept = .pct_ref, linetype = "dotdash") +
  geom_point(aes(y = N, color = THRES2), alpha = 0.3) +
  geom_line( aes(y = N, colour = THRES2, group = 1)) +
  theme_minimal()
 return(p)
}

 gg <- plot_timeline(us, us_ref2) 
 gg
```

@@

## Network Level Assessment for Brazil, Europe, and the United States

Based on the study dataset Fig. \@ref(fig:osndaily) provides an appreciation of the global developments.
This paper zooms in on comparing the situation in Brazil, Europe, and the United States.
Figure \@ref(fig:us-eu-br-timeline) shows the different network level responses to the regional/local air traffic constraints.

```{r}
us_timeline <- regional_timeline(ds, "US")
eu_timeline <- regional_timeline(ds, "EU")
br_timeline <- regional_timeline(ds, "BR")

trim_timeline <- function(.df){
  df <- .df %>% filter(DATE >= lubridate::ymd("2020-01-01"))
}

us_p <- plot_timeline(us_timeline[[1]] %>% trim_timeline(), us_timeline[[3]])
eu_p <- plot_timeline(eu_timeline[[1]] %>% trim_timeline(), eu_timeline[[3]])
br_p <- plot_timeline(br_timeline[[1]] %>% trim_timeline(), br_timeline[[3]])
```

(ref:us-eu-br-timeline) Timeline for United States, Europe, and Brazil

```{r us-eu-br-timeline, fig.cap="(ref:us-eu-br-timeline)"}
( (us_p + labs(x=NULL, y=NULL, subtitle = "United States")) + 
  (eu_p + labs(x=NULL, y=NULL, subtitle = "Europe"       )) + 
  (br_p + labs(x=NULL, y=NULL, subtitle = "Brazil"       ))   ) + 
  plot_layout(guides = "collect") &
  theme(legend.position='bottom')
```

The United States have recovered.

Europe has seen an initial recovery in summer 2020 with a decline towards the holiday season 2020/2021.
With April/May 2021 traffic increases again in Europe.

Brazil also experienced the sharp decline in traffic in March 2020.
Similar to the US, traffic in Brazil showed an initial continual increase. 
Traffic show an alternating pattern.

```
TODO: provide some more commentary
```

## Traffic Development - Connectivity
